Shannon-Inspired Information in the Clinical Use of Neural Signals Concerning Post-Comatose Patients

### 2023-02-01 ART_Noh/Noh.tex
Noh article added

\begin{document}
\title{Shannon-Inspired Information in the Clinical Use of Neural Signals Concerning Post-Comatose Patients}
\maketitle

\section*{Abstract}
\begin{quotation}
Post-comatose patients are classified as being in a~minimally conscious state when they have executive functions. Because traditional behavioral assessments may not capture signs of executive functions in post-comatose patients, clinicians look to localized brain activities in response to task instructions, such as imagining wiggling toes, to diagnose minimal consciousness. This paper critically assesses the assumption underlying such alternative methods: that brain activities are neural signals conveying information about minimal consciousness. Based on a~Shannon-inspired idea of information, I~distinguish between informational and engineering aspects of clinical tasks. The informational aspect concerns the conditional probability that, for example, given activity in the motor areas of the brain in response to task instructions, a~patient is imagining wiggling toes. The engineering aspect concerns efficient activation of the relevant brain areas in a~patient under the task conditions. This distinction shows that the current alternative methods are not informationally problematic, but are structurally ``ill-formed.'' For instance, the toe-imagery task requires the capacity to comprehend syntactically complex sentences, which can be dissociated from minimal consciousness. I~propose a~misrepresentation task, which tests the capacity to misconceptualize lukewarm water as melting wax, as a~supplement to the current alternative methods. This task is as informationally reliable as these methods, but is structurally ``well-formed,'' as it does not rely methodologically on prerequisites such as language comprehension.

\end{quotation}
\section*{Keyword}
\begin{quotation}
Post-comatose disorders of consciousness, Shannon's theory of information, Minimal consciousness, Mental action

\end{quotation}
\section{Introduction}
Minimally conscious post-comatose patients (henceforth, MCP patients) can have limited executive functions that allow for the patients to follow simple instructions or respond to certain stimulations. Signs of residual executive functions provide clinicians with strong evidence of eventual recovery of consciousness
%\label{ref:RNDc52VR2C0H1}(Naccache, 2018; Rohaut, Eliseyev and Claassen, 2019).
\parencites[][]{naccache_minimally_2018}[][]{rohaut_uncovering_2019}. %
 Thus, delayed recognition of such signs may lead to suboptimal clinical care 
%\label{ref:RND36iYhUl47e}(van Erp et al., 2019),
\parencite[van][]{van_erp_unexpected_2019}, %
 distorted prognostic figures for rehabilitation 
%\label{ref:RNDDiumYQqtvx}(Ansell, 1993),
\parencite[][]{ansell_slow--recover_1993}, %
 or ethical problems 
%\label{ref:RNDY4B7wdnTLX}(Peterson and Bayne, 2018; Noh, 2022).
\parencites[][]{peterson_post-comatose_2018}[][]{noh_behavioral_2022}. %
 Such signs, however, might not be captured by traditional behavioral assessments (e.g., eye tracking, speaking, etc.) because limited executive functions can be dissociated from the capacity to perform overt behaviors 
%\label{ref:RND1VSfdAw16s}(Teasdale and Jennett, 1974; Andrews et al., 1996).
\parencites[][]{teasdale_assessment_1974}[][]{andrews_misdiagnosis_1996}. %
 Although some traditional behavioral assessment methods can improve diagnostic accuracy 
%\label{ref:RNDkWoQFW7Rb7}(Schnakers et al., 2009),
\parencite[][]{schnakers_diagnostic_2009}, %
 differentiating MCP patients from patients in a~vegetative state remains challenging 
%\label{ref:RNDXmlWJMWjeY}(van Erp et al., 2015).
\parencite[van][]{van_erp_vegetative_2015}.%

@@

The aim of this paper is to critically assess whether NSAs satisfy this overarching assumption. Drawing on a~Shannon-inspired idea of information, I~distinguish between informational and engineering aspects of neural signal processing in the relevant studies. Shannon's
%\label{ref:RNDzWoTt1DLZu}(1948)
\parencite*[][]{} %
 theory of information ``defines the quantity of information in a~signal over the space of possibilities in a~given situation where a~sender and a~receiver are communicating'' 
%\label{ref:RNDESdIFerObH}(Noh, 2018, p.179).
\parencite[][p.179]{noh_no-report_2018}. %
 The primary concern of this paper is the communication channel between an MCP patient and a~clinician (or a~test-subject and an experimenter), where signal-vehicles are localized brain activities measured using fMRI or EEG. I~apply Shannon's theory to this communication channel in order to estimate the quantity of information about minimal consciousness possibly conveyed by the neural signal. Following Weaver 
%\label{ref:RND12k0aw3PZt}(1953, p.270),
\parencite*[][p.270]{weaver_recent_1953}, %
 I~further discuss whether the communication channel is designed to efficiently handle the assigned jobs. Notice that I~use the term ``Shannon-inspired idea of information'' rather than ``Shannon's theory of information'' because Shannon did not intend the theory to be applied to this sort of communication channel, and I~make certain assumptions in order to apply it here (see footnotes 2 and 3 for details of these assumptions).

@@

An essential difference between active and passive paradigms is that while the former requires a~subject to either perform a~mental-motor action or count occurrences of a~target, the latter concerns whether a~subject is able to experience stimulation without performing a~mental action. Nonetheless, NSAs in general appeal to localized brain activities, which can be captured either by fMRI or by EEG, as a~means of diagnosis of minimal consciousness.\footnote{ For the sake of simplicity, I~am going to ignore methodological differences in the use of fMRI and EEG, which are irrelevant to my argument.} Specifically, NSAs assume that the observed brain activities are neural signals conveying the message that the subjects under consideration are following instructions or paying attention to stimulations. To analyze the informational aspect is to assess whether the signals do convey the purported message.

Consider Cruse et al.'s
%\label{ref:RNDAh4jo6f2zc}(2011)
\parencite*[][]{cruse_bedside_2011} %
 toe-imagery task in order to analyze the informational aspect of active paradigm tasks. They instructed 16 behaviorally nonresponsive post-comatose patients (who were initially diagnosed with the vegetative state by traditional behavioral methods) and 12 healthy controls to imagine wiggling all of their toes on both feet and relaxing them without making any actual motor movement. Cruse et al. write that a~significant degree of activity was observed in a~relatively localized area of the medial premotor cortex of 3 patients and 9 controls. We can distinguish between the minimally conscious mental event, \textit{M}\textsubscript{1}, and a~set of nonconscious mental events, \textit{M}\textsubscript{2}, in relation to a~brain event, namely the activity in the medial premotor cortex, \textit{S}.\footnote{ Shannon's theory conceptualizes mutual information as the relation between events and a~subset of these events. In the body of the text, I~assume that mental events (\textit{M}\textsubscript{1} and \textit{M}\textsubscript{2}) are a~subset of a~brain event (\textit{S}). More specifically, I~assume that functional correlations between cognitive capacities and activities in the corresponding brain areas (e.g., relationships between particular types of motor actions and activities in the corresponding brain areas) lead to mutual information between the mental events and the brain event. Obviously, this assumption is controversial. Nonetheless, in the current paper, I~have made an attempt to demonstrate that such an information-theoretic assumption inevitably leads to a~clinically important conclusion. See 
%\label{ref:RNDi14ONyoLX3}(Li et al., 2022)
\parencite[][]{li_brain_2022} %
 for a~detailed discussion of the quantitative relationships between localized brain activities and information processing capacities. Kycia 
%\label{ref:RNDIqZJ434484}(2021)
\parencite*[][]{kycia_information_2021} %
 provides good clarification of fundamental issues concerning classical and quantum nature of information storing and processing in the brain.} To analyze the informational aspect for active paradigms is to compare \textit{M}\textsubscript{1} and \textit{M}\textsubscript{2}.\footnote{ Strictly speaking, \textit{M}\textsubscript{1} should be a~disjunction of the minimally conscious state and the fully conscious state (in Cruse et al., the healthy controls were fully conscious). Because the primary concern of NSAs is the minimally conscious state, however, we can safely ignore the information about the fully conscious state possibly conveyed by the neural signals.}

Mutual information between the brain event, \textit{S}, and a~set of mental events, \textit{M}, is:

$I(S:M)=H(M)H(MS)$.\footnote{ The idea of analyzing the informational aspect of active paradigm tasks originates from Noh
%\label{ref:RNDjBvUliVaOb}(2018, pp.181–185).
\parencite*[][pp.181–185]{noh_no-report_2018}.%
}

\textit{H} refers to Shannon entropy. To compare \textit{M}\textsubscript{1} and \textit{M}\textsubscript{2}, however, individual events need to be considered rather than the average. Because the toe-imagery task depends on the assumption that the detection of \textit{S} particularly reduces the uncertainty of the mental event that subjects are imagining wiggling toes, let's take \textit{M}\textsubscript{1} as referring to this mental event. The amount of uncertainty of \textit{M}\textsubscript{1} that \textit{S} reduces is:

$I(S:M_1)=I(M_1)I(M_1S)\log _2P(M_1)\log _2P(M_1S)\log _2\frac{P(M_1S)}{P(M_1)}$.

As $\log _2\frac{P(M_1S)}{P(M_1)}$ gets more and more positive, \textit{S} will represent greater accuracy about \textit{M}\textsubscript{1}. If $\log _2\frac{P(M_1S}){P(M_1})$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$, then the observation of \textit{S} provides clinicians with evidence that patients in the toe-imagery task who show activities in the medial premotor cortex are minimally conscious. If the latter is greater than, or equivalent to, the former, then \textit{S} alone cannot provide that evidence.

To estimate the value of $\log _2\frac{P(M_1S)}{P(M_1)}$, we need to understand the background conditions of the toe-imagery task. Notice that the patients in Cruse et al.'s
%\label{ref:RND2y1XLkJnMb}(2011)
\parencite*[][]{cruse_bedside_2011} %
 experiment were initially diagnosed with the vegetative state by traditional behavioral assessments. Given that the misdiagnosis rate of traditional behavioral assessments is 41\% 
%\label{ref:RNDQvnJPYKu8r}(Schnakers et al., 2009),
\parencite[][]{schnakers_diagnostic_2009}, %
 the unconditional probability that Cruse et al.'s patients were conscious, i.e., \textit{P}(\textit{M}\textsubscript{1}), is lower than the unconditional probability that they were in the vegetative state, i.e., \textit{P}(\textit{M}\textsubscript{2}). So, $\log _2\frac{P(M_1S}){P(M_1})$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$ if \textit{P}(\textit{M}\textsubscript{1}{\textbar}\textit{S}) is greater than, or equivalent to, \textit{P}(\textit{M}\textsubscript{2}{\textbar}\textit{S}).

\textit{P}(\textit{M}\textsubscript{1}{\textbar}\textit{S}) can be accounted for by the reversed strong correlation that holds between activities in motor areas and the corresponding mental-motor action. According to relevant experiments
%\label{ref:RND3XvevSKtpr}(e.g. Pfurtscheller and Neuper, 1997; Ehrsson, Geyer and Naito, 2003)
\parencites[e.g.,][]{pfurtscheller_motor_1997}[][]{ehrsson_imagery_2003} %
 in which healthy subjects were instructed to mentally act (e.g., they were instructed to imagine raising their left arms without making any actual motor movement), activities in the relevant motor areas were observed in the majority of the subjects. Provided that \textit{P}(activities in the relevant motor areas{\textbar}mental-motor action) is high, it seems that \textit{P}(\textit{M}\textsubscript{1}{\textbar}\textit{S}) is relatively high as well.

On the other hand, \textit{P}(\textit{M}\textsubscript{2}{\textbar}\textit{S}) concerns a~situation in which the medial premotor cortex activity in a~mental-motor action task, such as the toe-imagery task, is somehow automated (i.e., a~nonconscious response). Provided the strong functional relationship between the body surface and a~localized motor area (i.e., the well-established somatotopic map in motor areas), it seems possible to take the medial premotor cortex activity observed in the toe-imagery task as a~toe-related nonconscious mental event. In turn, to account for \textit{P}(\textit{M}\textsubscript{2}{\textbar}\textit{S}) would require finding a~toe-related nonconscious explanation (that does not involve volition, linguistic understanding of the verbal instruction, etc.) of why the activity occurred.

Nonetheless, no such explanation is available. Consider a~semantic priming effect, which refers to cases where automated activities in the somatotopy of the motor and premotor cortex are observed for a~few milliseconds when a~subject hears an action-word such as ``kick''
%\label{ref:RNDtVfb2YZ1ak}(Pulvermüller, 2005).
\parencite[][]{pulvermuller_brain_2005}. %
 Now, consider \textit{M}\textsubscript{2} as a~toe-related nonconscious mental event grounded by a~semantic priming effect. \textit{P}(\textit{M}\textsubscript{2}{\textbar}\textit{S}) in this sense, however, is very low for two reasons. First, the semantic priming effect has been observed when a~single word is given, but not when a~whole sentence, such as ``Imagine wiggling all of your toes,'' is given 
%\label{ref:RND5xoKfn4zNx}(Raposo et al., 2009).
\parencite[][]{raposo_modulation_2009}. %
 Second, the effect lasts for a~very short time, but the activity in Cruse et al.'s experiment persisted for more than a~few seconds. Because there is no alternative explanation, $\log _2\frac{P(M_1S}){P(M_1})$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$.

Following the same line of reasoning to analyze the informational aspect of passive paradigm tasks, let us consider King et al.'s
%\label{ref:RND2LXLkT9KBe}(2013)
\parencite*[][]{lule_probing_2013} %
 and Bekinschtein et al.'s 
%\label{ref:RNDmqV1uruVqo}(2009)
\parencite*[][]{schnakers_diagnostic_2009} %
 oddball tasks. The oddball task is designed to evaluate cerebral responses to violations of temporal regularities that are global across several seconds. Suppose that a~stream of repeated auditory events is given to a~healthy subject, with two types of novel events (i.e., oddballs) embedded in the stream. The two types are local oddballs, which consist of a~change in pitch within a~five-sound sequence (e.g., AAAAB), and global oddballs, which consist of a~change in an auditory sequence in a~fixed global context (e.g., AAAAB AAAAB AAAAB AAAAB AAAAA). Local oddballs typically lead to so-called frontal mismatch negativity, which is an automated (i.e., nonconscious) brain activity. In other words, frontal mismatch negativity is observed independently of whether a~subject pays attention to the given auditory stream or not. On the other hand, global oddballs generate the so-called P300b response, which can be detected only when a~subject is consciously paying attention to the given auditory stream. In King et al.'s 
%\label{ref:RNDslg878rAtY}(2013)
\parencite*[][]{lule_probing_2013} %
 experiment, a~global effect was found in 14\% of 70 vegetative state patients and 31\% of 65 minimally conscious state patients (and 52\% of 23 conscious controls with brain injuries). In Bekinschtein et al.'s 
%\label{ref:RNDBL3kglmU8Y}(2009)
\parencite*[][]{schnakers_diagnostic_2009} %
 experiment, a~global effect was found in none of 3 vegetative state patients and 3 of 4 minimally conscious state patients (and all of 11 healthy controls). But there was no significant difference between the patient group and the control group regarding the local effect in both experiments.

@@

Recall that in Cruse et al.'s
%\label{ref:RND5hclfQLLkX}(2011)
\parencite*[][]{cruse_bedside_2011} %
 toe-imagery task, a~strong correlation between activities in motor areas and the corresponding mental-motor action holds because \textit{P}(\textit{M}\textsubscript{1}{\textbar}\textit{S}) is relatively greater than \textit{P}(\textit{M}\textsubscript{2}{\textbar}\textit{S}).\footnote{ Notice that \textit{P}(\textit{M}\textsubscript{2}) is always greater than \textit{P}(\textit{M}\textsubscript{1}) in both active and passive paradigms because every MCP patient in the experiments was initially diagnosed with the vegetative state by traditional behavioral assessments. Specifically, because the misdiagnosis rate of traditional behavior assessments is 41\%, \textit{P}(\textit{M}\textsubscript{2}) is roughly 60\%.} So, we should compare \textit{P}(the subject is minimally conscious{\textbar}the P300b response) and \textit{P}(the subject is nonconscious{\textbar}the P300b response). According to the relevant studies, the P300b response requires working memory. Specifically, the P300b response is accounted for by the following hypotheses: Noticing global oddballs requires working memory and predictive coding 
%\label{ref:RNDXe3oyEzvUf}(Garrido et al., 2008; 2009);
\parencites*[][]{garrido_functional_2008}[][]{schnakers_diagnostic_2009}; %
 the P300b response is a~neural signature of postperceptual processing, such as working memory 
%\label{ref:RNDD9OoRT2KMc}(Cohen et al., 2020);
\parencite[][]{cohen_distinguishing_2020}; %
 and the P300b response is a~neural signature of working memory–guided categorization processes 
%\label{ref:RND1OFfKua7Q6}(Rac-Lubashevsky and Kessler, 2019).
\parencite[][]{rac-lubashevsky_revisiting_2019}. %
 If there is no alternative explanation of the P300b response, and if working memory is a~sign of minimal consciousness 
%\label{ref:RNDkNGafGH8qa}(Ansell, 1993; Bekinschtein et al., 2009; King et al., 2013),
\parencites[][]{ansell_slow--recover_1993}[][]{bekinschtein_neural_2009}[][]{king_single-trial_2013}, %
 then \textit{P}(the subject is minimally conscious{\textbar}the P300b response) is greater than \textit{P}(the subject is nonconscious{\textbar}the P300b response).

@@

The toe-imagery task is structurally ``ill-formed'' because it depends on verbal linguistic instructions. In Cruse et al.'s experiment, it might be that some MCP patients did not show the purported brain activities in the toe-imagery task, not because they were not minimally conscious, but because they could not clearly comprehend the instructions and conceived the visual motor imagery instead. According to Kwiatkowska et al.
%\label{ref:RNDZNjyq1zhPj}(2019),
\parencite*[][]{vassilieva_automated_2019}, %
 34\% of 50 minimally conscious post-comatose patients could not build and had difficulties in reading syntactically complex sentences. Most importantly, the instructions of the toe-imagery task consist of sentences that are syntactically far more complex than those in Kwiatkowska et al.'s experiment. Moreover, in general, MCP patients' brain responses to heard words are weaker in terms of power than those of healthy controls 
%\label{ref:RNDEz2DkUdcqj}(Nigri et al., 2017).
\parencite[][]{nigri_neural_2017}. %
 In a~nutshell, the capacity to comprehend syntactically complex spoken sentences can be dissociated from minimal consciousness. Thus, the toe-imagery task structurally permits the possibility of false negatives.

>>>>

\begin{artengenv}{Hyungrae Noh}
	{Shannon-inspired information in the clinical use of neural signals concerning post-comatose patients}
	{Shannon-inspired information in the clinical use of neural signals\ldots}
	{Shannon-inspired information in the clinical use of neural signals concerning post-comatose patients}
	{Sunchon National Unviersity}
	{Post-comatose patients are classified as being in a~minimally conscious state when they have executive functions. Because traditional behavioral assessments may not capture signs of executive functions in post-comatose patients, clinicians look to localized brain activities in response to task instructions, such as imagining wiggling toes, to diagnose minimal consciousness. This paper critically assesses the assumption underlying such alternative methods: that brain activities are neural signals conveying information about minimal consciousness. Based on a~Shannon-inspired idea of information, I~distinguish between informational and engineering aspects of clinical tasks. The informational aspect concerns the conditional probability that, for example, given activity in the motor areas of the brain in response to task instructions, a~patient is imagining wiggling toes. The engineering aspect concerns efficient activation of the relevant brain areas in a~patient under the task conditions. This distinction shows that the current alternative methods are not informationally problematic, but are structurally ``ill-formed.'' For instance, the toe-imagery task requires the capacity to comprehend syntactically complex sentences, which can be dissociated from minimal consciousness. I~propose a~misrepresentation task, which tests the capacity to misconceptualize lukewarm water as melting wax, as a~supplement to the current alternative methods. This task is as informationally reliable as these methods, but is structurally ``well-formed,'' as it does not rely methodologically on prerequisites such as language comprehension.
	}
	{post-comatose disorders of consciousness, Shannon's theory of information, minimal consciousness, mental action.}

\section{Introduction}
Minimally conscious post-comatose patients (henceforth, MCP patients) can have limited executive functions that allow for the patients to follow simple instructions or respond to certain stimulations. Signs of residual executive functions provide clinicians with strong evidence of eventual recovery of consciousness
%\label{ref:RNDc52VR2C0H1}(Naccache, 2018; Rohaut, Eliseyev and Claassen, 2019).
\parencites[][]{naccache_minimally_2018}[][]{rohaut_uncovering_2019}. %
 Thus, delayed recognition of such signs may lead to suboptimal clinical care 
%\label{ref:RND36iYhUl47e}(van Erp et al., 2019),
\parencite[van][]{van_erp_unexpected_2019}, %
 distorted prognostic figures for rehabilitation 
%\label{ref:RNDDiumYQqtvx}(Ansell, 1993),
\parencite[][]{ansell_slow--recover_1993}, %
 or ethical problems 
%\label{ref:RNDY4B7wdnTLX}(Peterson and Bayne, 2018; Noh, 2022).
\parencites[][]{peterson_post-comatose_2018}[][]{noh_behavioral_2022}. %
 Such signs, however, might not be captured by traditional behavioral assessments (e.g., eye tracking, speaking, etc.) because limited executive functions can be dissociated from the capacity to perform overt behaviors 
%\label{ref:RND1VSfdAw16s}(Teasdale and Jennett, 1974; Andrews et al., 1996).
\parencites[][]{teasdale_assessment_1974}[][]{andrews_misdiagnosis_1996}. %
 Although some traditional behavioral assessment methods can improve diagnostic accuracy 
%\label{ref:RNDkWoQFW7Rb7}(Schnakers et al., 2009),
\parencite[][]{schnakers_diagnostic_2009}, %
 differentiating MCP patients from patients in a~vegetative state remains challenging 
%\label{ref:RNDXmlWJMWjeY}(van Erp et al., 2015).
\parencite[van][]{van_erp_vegetative_2015}.%

@@

The aim of this paper is to critically assess whether NSAs satisfy this overarching assumption. Drawing on a~Shannon-inspired idea of information, I~distinguish between informational and engineering aspects of neural signal processing in the relevant studies. Shannon's
%\label{ref:RNDzWoTt1DLZu}(1948)
\parencite*[][]{shannon_mathematical_1948} %
 theory of information ``defines the quantity of information in a~signal over the space of possibilities in a~given situation where a~sender and a~receiver are communicating'' 
%\label{ref:RNDESdIFerObH}(Noh, 2018, p.179).
\parencite[][p.179]{noh_no-report_2018}. %
 The primary concern of this paper is the communication channel between an MCP patient and a~clinician (or a~test-subject and an experimenter), where signal-vehicles are localized brain activities measured using fMRI or EEG. I~apply Shannon's theory to this communication channel in order to estimate the quantity of information about minimal consciousness possibly conveyed by the neural signal. Following Weaver 
%\label{ref:RND12k0aw3PZt}(1953, p.270),
\parencite*[][p.270]{weaver_recent_1953}, %
 I~further discuss whether the communication channel is designed to efficiently handle the assigned jobs. Notice that I~use the term ``Shannon-inspired idea of information'' rather than ``Shannon's theory of information'' because Shannon did not intend the theory to be applied to this sort of communication channel, and I~make certain assumptions in order to apply it here (see footnotes 2 and 3 for details of these assumptions).

@@

An essential difference between active and passive paradigms is that while the former requires a~subject to either perform a~mental-motor action or count occurrences of a~target, the latter concerns whether a~subject is able to experience stimulation without performing a~mental action. Nonetheless, NSAs in general appeal to localized brain activities, which can be captured either by fMRI or by EEG, as a~means of diagnosis of minimal consciousness.\footnote{For the sake of simplicity, I~am going to ignore methodological differences in the use of fMRI and EEG, which are irrelevant to my argument.} Specifically, NSAs assume that the observed brain activities are neural signals conveying the message that the subjects under consideration are following instructions or paying attention to stimulations. To analyze the informational aspect is to assess whether the signals do convey the purported message.

Consider Cruse et al.'s
%\label{ref:RNDAh4jo6f2zc}(2011)
\parencite*[][]{cruse_bedside_2011} %
 toe-imagery task in order to analyze the informational aspect of active paradigm tasks. They instructed 16 behaviorally nonresponsive post-comatose patients (who were initially diagnosed with the vegetative state by traditional behavioral methods) and 12 healthy controls to imagine wiggling all of their toes on both feet and relaxing them without making any actual motor movement. Cruse et al. write that a~significant degree of activity was observed in a~relatively localized area of the medial premotor cortex of 3 patients and 9 controls. We can distinguish between the minimally conscious mental event, $M_1$, and a~set of nonconscious mental events, $M_2$, in relation to a~brain event, namely the activity in the medial premotor cortex, $S$.\footnote{Shannon's theory conceptualizes mutual information as the relation between events and a~subset of these events. In the body of the text, I~assume that mental events ($M_1$ and $M_2$) are a~subset of a~brain event ($S$). More specifically, I~assume that functional correlations between cognitive capacities and activities in the corresponding brain areas (e.g., relationships between particular types of motor actions and activities in the corresponding brain areas) lead to mutual information between the mental events and the brain event. Obviously, this assumption is controversial. Nonetheless, in the current paper, I~have made an attempt to demonstrate that such an information-theoretic assumption inevitably leads to a~clinically important conclusion. See 
%\label{ref:RNDi14ONyoLX3}(Li et al., 2022)
\parencite[][]{li_brain_2022} %
 for a~detailed discussion of the quantitative relationships between localized brain activities and information processing capacities. Kycia 
%\label{ref:RNDIqZJ434484}(2021)
\parencite*[][]{kycia_information_2021} %
 provides good clarification of fundamental issues concerning classical and quantum nature of information storing and processing in the brain.} To analyze the informational aspect for active paradigms is to compare $M_1$ and $M_2$.\footnote{Strictly speaking, $M_1$ should be a~disjunction of the minimally conscious state and the fully conscious state (in Cruse et al., the healthy controls were fully conscious). Because the primary concern of NSAs is the minimally conscious state, however, we can safely ignore the information about the fully conscious state possibly conveyed by the neural signals.}

Mutual information between the brain event, $S$, and a~set of mental events, $M$, is:
$$I(S:M)=H(M)H(MS).\footnote{The idea of analyzing the informational aspect of active paradigm tasks originates from Noh
%\label{ref:RNDjBvUliVaOb}(2018, pp.181–185).
\parencite*[][pp.181–185]{noh_no-report_2018}.%
}$$
\textit{H} refers to Shannon entropy. To compare $M_1$ and $M_2$, however, individual events need to be considered rather than the average. Because the toe-imagery task depends on the assumption that the detection of $S$ particularly reduces the uncertainty of the mental event that subjects are imagining wiggling toes, let's take $M_1$ as referring to this mental event. The amount of uncertainty of $M_1$ that $S$ reduces is:
$$I(S:M_1)=I(M_1)I(M_1S)\log _2P(M_1)\log _2P(M_1S)\log _2\frac{P(M_1S)}{P(M_1)}.$$
As $\log _2\frac{P(M_1S)}{P(M_1)}$ gets more and more positive, $S$ will represent greater accuracy about $M_1$. If $\log _2\frac{P(M_1S}{P(M_1)}$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$, then the observation of $S$ provides clinicians with evidence that patients in the toe-imagery task who show activities in the medial premotor cortex are minimally conscious. If the latter is greater than, or equivalent to, the former, then $S$ alone cannot provide that evidence.

To estimate the value of $\log _2\frac{P(M_1S)}{P(M_1)}$, we need to understand the background conditions of the toe-imagery task. Notice that the patients in Cruse et al.'s
%\label{ref:RND2y1XLkJnMb}(2011)
\parencite*[][]{cruse_bedside_2011} %
 experiment were initially diagnosed with the vegetative state by traditional behavioral assessments. Given that the misdiagnosis rate of traditional behavioral assessments is 41\% 
%\label{ref:RNDQvnJPYKu8r}(Schnakers et al., 2009),
\parencite[][]{schnakers_diagnostic_2009}, %
 the unconditional probability that Cruse et al.'s patients were conscious, i.e., $P(M_1)$, is lower than the unconditional probability that they were in the vegetative state, i.e., $P(M_2)$. So, $\log _2\frac{P(M_1S}{P(M_1)}$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$ if $P(M_1 | S)$ is greater than, or equivalent to, $P(M_2 | S)$.

$P(M_1 | S)$ can be accounted for by the reversed strong correlation that holds between activities in motor areas and the corresponding mental-motor action. According to relevant experiments
%\label{ref:RND3XvevSKtpr}(e.g. Pfurtscheller and Neuper, 1997; Ehrsson, Geyer and Naito, 2003)
\parencites[e.g.,][]{pfurtscheller_motor_1997}[][]{ehrsson_imagery_2003} %
 in which healthy subjects were instructed to mentally act (e.g., they were instructed to imagine raising their left arms without making any actual motor movement), activities in the relevant motor areas were observed in the majority of the subjects. Provided that $P$ (activities in the relevant motor areas{\textbar}mental-motor action) is high, it seems that $P(M_1 | S)$ is relatively high as well.

On the other hand, $P(M_2 | S)$ concerns a~situation in which the medial premotor cortex activity in a~mental-motor action task, such as the toe-imagery task, is somehow automated (i.e., a~nonconscious response). Provided the strong functional relationship between the body surface and a~localized motor area (i.e., the well-established somatotopic map in motor areas), it seems possible to take the medial premotor cortex activity observed in the toe-imagery task as a~toe-related nonconscious mental event. In turn, to account for $P(M_2 | S)$ would require finding a~toe-related nonconscious explanation (that does not involve volition, linguistic understanding of the verbal instruction, etc.) of why the activity occurred.

Nonetheless, no such explanation is available. Consider a~semantic priming effect, which refers to cases where automated activities in the somatotopy of the motor and premotor cortex are observed for a~few milliseconds when a~subject hears an action-word such as ``kick''
%\label{ref:RNDtVfb2YZ1ak}(Pulvermüller, 2005).
\parencite[][]{pulvermuller_brain_2005}. %
 Now, consider $M_2$ as a~toe-related nonconscious mental event grounded by a~semantic priming effect. $P(M_2 | S)$ in this sense, however, is very low for two reasons. First, the semantic priming effect has been observed when a~single word is given, but not when a~whole sentence, such as ``Imagine wiggling all of your toes,'' is given 
%\label{ref:RND5xoKfn4zNx}(Raposo et al., 2009).
\parencite[][]{raposo_modulation_2009}. %
 Second, the effect lasts for a~very short time, but the activity in Cruse et al.'s experiment persisted for more than a~few seconds. Because there is no alternative explanation, $\log _2\frac{P(M_1S}{P(M_1)}$ is greater than $\log _2\frac{P(M_2S)}{P(M_2)}$.

Following the same line of reasoning to analyze the informational aspect of passive paradigm tasks, let us consider King et al.'s
%\label{ref:RND2LXLkT9KBe}(2013)
\parencite*[][]{king_single-trial_2013} %
 and Bekinschtein et al.'s 
%\label{ref:RNDmqV1uruVqo}(2009)
\parencite*[][]{bekinschtein_neural_2009} %
 oddball tasks. The oddball task is designed to evaluate cerebral responses to violations of temporal regularities that are global across several seconds. Suppose that a~stream of repeated auditory events is given to a~healthy subject, with two types of novel events (i.e., oddballs) embedded in the stream. The two types are local oddballs, which consist of a~change in pitch within a~five-sound sequence (e.g., AAAAB), and global oddballs, which consist of a~change in an auditory sequence in a~fixed global context (e.g., AAAAB AAAAB AAAAB AAAAB AAAAA). Local oddballs typically lead to so-called frontal mismatch negativity, which is an automated (i.e., nonconscious) brain activity. In other words, frontal mismatch negativity is observed independently of whether a~subject pays attention to the given auditory stream or not. On the other hand, global oddballs generate the so-called P300b response, which can be detected only when a~subject is consciously paying attention to the given auditory stream. In King et al.'s 
%\label{ref:RNDslg878rAtY}(2013)
\parencite*[][]{king_single-trial_2013} %
 experiment, a~global effect was found in 14\% of 70 vegetative state patients and 31\% of 65 minimally conscious state patients (and 52\% of 23 conscious controls with brain injuries). In Bekinschtein et al.'s 
%\label{ref:RNDBL3kglmU8Y}(2009)
\parencite*[][]{bekinschtein_neural_2009} %
 experiment, a~global effect was found in none of 3 vegetative state patients and 3 of 4 minimally conscious state patients (and all of 11 healthy controls). But there was no significant difference between the patient group and the control group regarding the local effect in both experiments.

@@

Recall that in Cruse et al.'s
%\label{ref:RND5hclfQLLkX}(2011)
\parencite*[][]{cruse_bedside_2011} %
 toe-imagery task, a~strong correlation between activities in motor areas and the corresponding mental-motor action holds because $P(M_1 | S)$ is relatively greater than $P(M_2 | S)$.\footnote{Notice that $P(M_2)$ is always greater than $P(M_1)$ in both active and passive paradigms because every MCP patient in the experiments was initially diagnosed with the vegetative state by traditional behavioral assessments. Specifically, because the misdiagnosis rate of traditional behavior assessments is 41\%, $P(M_2)$ is roughly 60\%.} So, we should compare $P$ (the subject is minimally conscious{\textbar}the P300b response) and $P$ (the subject is nonconscious{\textbar}the P300b response). According to the relevant studies, the P300b response requires working memory. Specifically, the P300b response is accounted for by the following hypotheses: Noticing global oddballs requires working memory and predictive coding 
%\label{ref:RNDXe3oyEzvUf}(Garrido et al., 2008; 2009);
\parencites[][]{garrido_functional_2008}[][]{garrido_repetition_2009}; %
 the P300b response is a~neural signature of postperceptual processing, such as working memory 
%\label{ref:RNDD9OoRT2KMc}(Cohen et al., 2020);
\parencite[][]{cohen_distinguishing_2020}; %
 and the P300b response is a~neural signature of working memory–guided categorization processes 
%\label{ref:RND1OFfKua7Q6}(Rac-Lubashevsky and Kessler, 2019).
\parencite[][]{rac-lubashevsky_revisiting_2019}. %
 If there is no alternative explanation of the P300b response, and if working memory is a~sign of minimal consciousness 
%\label{ref:RNDkNGafGH8qa}(Ansell, 1993; Bekinschtein et al., 2009; King et al., 2013),
\parencites[][]{ansell_slow--recover_1993}[][]{bekinschtein_neural_2009}[][]{king_single-trial_2013}, %
 then $P$(the subject is minimally conscious{\textbar}the P300b response) is greater than $P$ (the subject is nonconscious{\textbar}the P300b response).

@@

The toe-imagery task is structurally ``ill-formed'' because it depends on verbal linguistic instructions. In Cruse et al.'s experiment, it might be that some MCP patients did not show the purported brain activities in the toe-imagery task, not because they were not minimally conscious, but because they could not clearly comprehend the instructions and conceived the visual motor imagery instead. According to Kwiatkowska et al.
%\label{ref:RNDZNjyq1zhPj}(2019),
\parencite*[][]{kwiatkowska_post-comatose_2019}, %
 34\% of 50 minimally conscious post-comatose patients could not build and had difficulties in reading syntactically complex sentences. Most importantly, the instructions of the toe-imagery task consist of sentences that are syntactically far more complex than those in Kwiatkowska et al.'s experiment. Moreover, in general, MCP patients' brain responses to heard words are weaker in terms of power than those of healthy controls 
%\label{ref:RNDEz2DkUdcqj}(Nigri et al., 2017).
\parencite[][]{nigri_neural_2017}. %
 In a~nutshell, the capacity to comprehend syntactically complex spoken sentences can be dissociated from minimal consciousness. Thus, the toe-imagery task structurally permits the possibility of false negatives.
